---
title: 'Assignment 09: Data Scraping'
author: "Yingchi Cheung"
output:
  pdf_document: default
  html_document:
    df_print: paged
geometry: margin=2.54cm
editor_options:
  chunk_output_type: inline
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A09_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1
getwd()
library(tidyverse)
library(lubridate)
library(rvest)

mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom")
theme_set(mytheme)
```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2021 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2
webpage <- read_html('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2021')

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings), with the first value being "27.6400".

```{r scrape.the.data}
#3
water.system.name <- webpage %>% html_nodes('div+ table tr:nth-child(1) td:nth-child(2)') %>% html_text()
water.system.name

pswid <- webpage %>% html_nodes('td tr:nth-child(1) td:nth-child(5)') %>% html_text()
pswid

ownership <- webpage %>% html_nodes('div+ table tr:nth-child(2) td:nth-child(4)') %>% html_text()
ownership

max.withdrawals.mgd <- webpage %>% html_nodes('th~ td+ td') %>% html_text()
max.withdrawals.mgd

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc...

5. Create a line plot of the maximum daily withdrawals across the months for 2021

```{r create.a.dataframe.from.scraped.data}
#4
Month <- c(1,5,9,2,6,10,3,7,11,4,8,12)

Withdrawals <-
  data.frame("Max_Withdrawals_mgd" = as.numeric(max.withdrawals.mgd),
             "Month" = Month,
             "Year" = rep(2021,12)) %>%
  mutate(Water_System_Name = !!water.system.name, PWSID = !!pswid, Ownership= !!ownership, Date = my(paste(Month,"-", Year)))

#5
ggplot(Withdrawals,aes(x=Date,y=Max_Withdrawals_mgd)) + 
  geom_line() + 
  labs(title = paste("2021 Maxium Daily Water Withdrawals for", water.system.name),
       subtitle = pswid,
       y="Maximum Withdrawal (mgd)",
       x="Date")

```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site (pwsid) scraped**.

```{r construct.a.scraping.function}
#6.
scrape.it <- function(The_pwsid, The_year){
  
The_webpage <- read_html(paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=', The_pwsid, '&year=', The_year))
The_water_system_name<-'div+ table tr:nth-child(1) td:nth-child(2)'
The_pswid<-'td tr:nth-child(1) td:nth-child(5)'
The_ownership<-'div+ table tr:nth-child(2) td:nth-child(4)'
The_max_withdrawals<-'th~ td+ td'


water.system.name <- The_webpage %>% html_nodes(The_water_system_name) %>% html_text()

pswid <- The_webpage %>% html_nodes(The_pswid) %>% html_text()

ownership <- The_webpage %>% html_nodes(The_ownership) %>% html_text()

max.withdrawals.mgd <- The_webpage %>% html_nodes(The_max_withdrawals) %>% html_text()

Month <- c(1,5,9,2,6,10,3,7,11,4,8,12)

Withdrawals <-
  data.frame("Max_Withdrawals_mgd" = as.numeric(max.withdrawals.mgd),
             "Month" = Month,
             "Year" = rep(The_year, 12)) %>%
  mutate(Water_System_Name = !!water.system.name, PWSID = !!pswid, Ownership= !!ownership, Date = my(paste(Month,"-", Year)))

return(Withdrawals)
}
```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
Durham <- scrape.it('03-32-010', '2015')
ggplot(Durham, aes(x = Date,y = Max_Withdrawals_mgd)) + 
  geom_line() + 
  labs(title = paste("2015 Maxium Daily Water Withdrawals for", water.system.name),
       subtitle = pswid,
       y="Maximum Withdrawal (mgd)",
       x="Date")

```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares Asheville's to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data}
#8
Asheville <- scrape.it('01-11-010','2015')
Both <- rbind(Durham, Asheville)

ggplot(Both, aes(x = Date,y = Max_Withdrawals_mgd, color = Water_System_Name)) + 
  geom_line() + 
  labs(title = paste("2015 Maxium Daily Water Withdrawals for Druham and Asheville"),
       subtitle = pswid,
       y="Maximum Withdrawal (mgd)",
       x="Date")

```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

>TIP: See Section 3.2 in the "09_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bindrows()` to combine the dataframes into a single one. 

```{r fetch.and.plot.Asheville.multiyear.data}
#9
years = rep(2010:2019)
A_pwsid = '01-11-010'

withdrawal_A <- lapply(X = years,
                  FUN = scrape.it,
                  The_pwsid = A_pwsid)

A2 <- bind_rows(withdrawal_A)

ggplot(A2, aes(x = Date, y = Max_Withdrawals_mgd)) + 
  geom_line() + 
  geom_smooth(method="loess",se=FALSE) +
  labs(title = paste("Maxium Daily Water Withdrawals for Asheville"),
       subtitle = "2010-2019",
       y="Maximum Withdrawal (mgd)",
       x="Date")
```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
> Asheville does have a trend in water usage over time. We can see there's an increasing trend line. Meaning that water usage increased from 2010 to 2019.