---
title: "Assignment 5: Data Visualization"
author: "Yingchi Cheung"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
#This code chunk will tidy your knit PDF files, wrapping long code lines
#For it to work, the "formatR" package needs to be installed

#install.packages('formatR')
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.

The completed exercise is due on Friday, Oct 21th @ 5:00pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse, 
lubridate, & cowplot packages. Upload the NTL-LTER processed data files for  
nutrients and chemistry/physics for Peter and Paul Lakes (use the tidy 
[`NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv`] version) and 
the processed data file for the Niwot Ridge litter dataset (use the
[`NEON_NIWO_Litter_mass_trap_Processed.csv`] version). 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r}
#1 
getwd()
library(tidyverse)
library(lubridate)
library(colormap)
NTLLTER <- read.csv("./Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv", stringsAsFactors = TRUE)
NEON_NIWO <- read.csv("./Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv", stringsAsFactors = TRUE)

#2 
NTLLTER$sampledate <- as.Date(NTLLTER$sampledate, format = "%Y-%m-%d")
class(NTLLTER$sampledate)
NEON_NIWO$collectDate <- as.Date(NEON_NIWO$collectDate, format = "%Y-%m-%d")
class(NEON_NIWO$collectDate)
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r}
#3
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom")
theme_set(mytheme)
```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best 
practices for data visualization. Ensure your theme, color palettes, axes, 
and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus (`tp_ug`) by phosphate (`po4`), with 
separate aesthetics for Peter and Paul lakes. Add a line of best fit and 
color it black. Adjust your axes to hide extreme values (hint: change the 
limits using `xlim()` and/or `ylim()`).

```{r}
#4
PvsPO4 <- 
  ggplot(NTLLTER, aes(x = po4, y = tp_ug)) +
  geom_point(aes(color = lakename)) +
  xlim(0, 50) +
  geom_smooth(method = lm, color = "black")+
  ylab("TP (micrograms per liter)")+
  xlab("PO4 (micrograms per liter)")
print(PvsPO4)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and 
(c) TN, with month as the x axis and lake as a color aesthetic. Then, create 
a cowplot that combines the three graphs. Make sure that only one legend is 
present and that graph axes are aligned. 

Tip: R has a build in variable called `month.abb` that returns a list of months;
see <https://r-lang.com/month-abb-in-r-with-example>

```{r}
#5
NTLLTER$month <- factor(NTLLTER$month, levels=c(1:12))
library(cowplot)
mnth <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)
month.abb[mnth]

Temperature <-
  ggplot(NTLLTER, aes(x = month, y = temperature_C)) +
  geom_boxplot(aes(color = lakename, alpha = 0.5)) +
  scale_x_discrete(drop=FALSE)+
  ggtitle("Temperature")+
  ylab("°C")+
  xlab("Month")
print(Temperature)

TP <-
  ggplot(NTLLTER, aes(x = month, y = tp_ug)) +
  geom_boxplot(aes(color = lakename, alpha = 0.5)) +
  scale_x_discrete(drop=FALSE)+
  labs(title="TP", y = "Micrograms Per Liter", x = "Month")
print(TP)

TN <-
  ggplot(NTLLTER, aes(x = month, y = tn_ug)) +
  geom_boxplot(aes(color = lakename, alpha = 0.5)) +
  scale_x_discrete(drop=FALSE)+
  labs(title="TN", y = "Micrograms Per Liter", x = "Month")
print(TN)
```
```{r, fig.height = 8, fig.width = 5}
plot_grid(Temperature + theme(legend.position = "none"), TP + theme(legend.position = "none"), TN, ncol = 1, align = "v", axis = 'l', rel_heights = c(1, 1, 1.2))

```

Question: What do you observe about the variables of interest over seasons and between lakes? 

> Answer: Peter Lake has a higher temperature than Paul Lake in November. But, the temperatures of both lakes in the rest of the months are not significantly different from others. Peter Lake's TP amount is likely to be higher than Paul Lake's in July and August. Overall, Peter Lack has higher outliers than Paul Lake, but not in March. In August, there is likely to be a higher amount of TN in Peter Lake than in Paul Lake. Peter Lack has higher outliers than Paul Lake throughout all year.

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the 
"Needles" functional group. Plot the dry mass of needle litter by date and 
separate by NLCD class with a color aesthetic. (no need to adjust the name 
of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into 
three facets rather than separated by color. 

```{r}
#6
Needles <- filter(NEON_NIWO, functionalGroup == "Needles")
litter_color <-
  ggplot(Needles, aes(x = collectDate, y = dryMass)) +
  geom_point(aes(color = nlcdClass), alpha = 0.5)
print(litter_color)
  
#7
litter_facets <-
  ggplot(Needles, aes(x = collectDate, y = dryMass)) +
  geom_point() +
   facet_wrap(vars(nlcdClass), nrow = 3)
print(litter_facets)

```
Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: The plot created in question 7 is more effective. The separation into three facets looks clearer than the color separation in question 6. In question 7, We can easily compare them by year and NLCD classes because they are not crowded together. But in question 6, even though they are categorized by year, it is still hard to tell these NLCD classes apart, let alone compare them.